---
title: "P8105_hw2_xh2470"
author: "Xueqing Huang"
output: github_document
---

```{r setup}
library(tidyverse)
library(readxl)
options(warn = -1)
```

# Problem 1

Read and clean the Mr.Trash Wheel sheet.

```{r}
trash_data = read_excel("./data/Trash-Wheel-Collection-Totals-7-2021-sept.xlsx", sheet = "Mr. Trash Wheel", range = "A2:N408") %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(sports_balls = round(sports_balls, digits = 0))
```

Read and clean precipitation data for 2018 and 2019, and combine them.

```{r}
data_2018 = read_excel("./data/Trash-Wheel-Collection-Totals-7-2021-sept.xlsx", sheet = "2018 Precipitation", range = "A2:B14") %>% 
  janitor::clean_names() %>%
  drop_na(total) %>% 
  mutate(year = "2018")

data_2019 = read_excel("./data/Trash-Wheel-Collection-Totals-7-2021-sept.xlsx", sheet = "2019 Precipitation", range = "A2:B14") %>% 
  janitor::clean_names() %>%
  drop_na(total) %>% 
  mutate(year = "2019")

data_combine = 
  bind_rows(data_2018, data_2019) %>% 
  mutate(month = month.name[month]) %>% 
  relocate(year, month)
```

Write a paragraph of these data.

For Mr.Trash Wheel dataset, there are **`r nrow(trash_data)`** rows and **`r ncol(trash_data)`** columns. The number of observations is **`r nrow(trash_data)`**. Variables contain _`r names(trash_data)`_.  Examples of key variables are:
`r knitr::kable(head(select(trash_data, dumpster, month, year, weight_tons, plastic_bottles)))`

For the combined precipitation dataset of 2018 and 2019, there are **`r nrow(data_combine)`** rows and **`r ncol(data_combine)`** columns. The number of observations is **`r nrow(data_combine)`**. Variables contain _`r names(data_combine)`_. Examples of key variables are:
`r knitr::kable(head(data_combine))`

The total precipitation in 2018 is **`r sum(pull(filter(data_combine, year == 2018), total))`**. The median number of sports balls in a dumpster in 2019 is **`r median(pull(filter(trash_data, year == 2019),sports_balls))`**.


# Problem 2

First, clean the data in pols-month.csv. 

```{r}
pols_month = read_csv("./data/pols-month.csv") %>% 
  janitor::clean_names() %>% 
  separate(mon, into = c("year", "month", "day")) %>% 
  mutate(month = month.name[as.numeric(month)]) %>% 
  mutate(president = ifelse(prez_gop == 1, "gop", "dem")) %>% 
  select(-day, -prez_gop, -prez_dem) %>% 
  arrange(year, month)
``` 

Second, clean the data in snp.csv.

```{r}
snp = read_csv("./data/snp.csv") %>% 
  janitor::clean_names() %>% 
  separate(date, into = c("month", "day", "year")) %>% 
  mutate(year = paste0(ifelse(as.numeric(year) <= 15, "20", "19"), year)) %>% 
  mutate(month = month.name[as.numeric(month)]) %>%   
  select(-day) %>% 
  relocate(year) %>% 
  arrange(year, month)
```

Third, tidy the unemployment data.

```{r}
unemployment = read_csv("./data/unemployment.csv") %>% 
  janitor::clean_names() %>%
  rename(January = jan, February = feb, March = mar, April = apr, May = may, June = jun, July = jul, August = aug, September = sep, October = oct, November = nov, December = dec) %>% 
  pivot_longer(
    January:December, 
    names_to = "month",
    values_to = "percentage"
  ) %>% 
  mutate(year = as.character(year)) %>% 
  arrange(year, month)
```

Fourth, join the datasets.

```{r}
# Join dataset pols_month and snp first.
join_a = left_join(pols_month, snp, by = c("year", "month")) 

# Then join dataset unemployment
join = left_join(join_a, unemployment, by = c("year", "month"))
```

Write a paragraph of these data.

The dataset `pols_month` gives information about the number of democratic and republican politicians, which contain **`r nrow(pols_month)`** observations and **`r ncol(pols_month)`** variables. Variables contain _`r names(pols_month)`_. The time is from `r min(as.numeric(pull(pols_month, year)))` to `r max(as.numeric(pull(pols_month, year)))`, and the range is `r max(as.numeric(pull(pols_month, year))) - min(as.numeric(pull(pols_month, year)))`.
The dataset `snp` reveals information about Standard & Poor’s stock market index (S&P), which contain **`r nrow(snp)`** observations and **`r ncol(snp)`** variables. Variables contain _`r names(snp)`_. The time is from `r min(as.numeric(pull(snp, year)))` to `r max(as.numeric(pull(snp, year)))`, and the range is `r max(as.numeric(pull(snp, year))) - min(as.numeric(pull(snp, year)))`.
The dataset `unemployment` shows percentage of unemployment in a given time, which contain **`r nrow(unemployment)`** observations and **`r ncol(unemployment)`** variables. Variables contain _`r names(unemployment)`_. The time is from `r min(as.numeric(pull(unemployment, year)))` to `r max(as.numeric(pull(unemployment, year)))`, and the range is `r max(as.numeric(pull(unemployment, year))) - min(as.numeric(pull(unemployment, year)))`.
The dataset `join` combines the above three datasets, and contains **`r nrow(join)`** observations and **`r ncol(join)`** variables.The time is from `r min(as.numeric(pull(join, year)))` to `r max(as.numeric(pull(join, year)))`, and the range is `r max(as.numeric(pull(join, year))) - min(as.numeric(pull(join, year)))`.
Examples of key variables are:
`r knitr::kable(head(join))`




