---
title: "P8105_hw2_xh2470"
author: "Xueqing Huang xh2470"
output: github_document
---

```{r setup}
library(tidyverse)
library(readxl)
options(warn = -1)
```

# Problem 1

Read and clean the Mr.Trash Wheel sheet.

```{r}
trash_data = read_excel("./data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", sheet = "Mr. Trash Wheel", range = "A2:N535") %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>%  # omit rows that do not include dumpster-specific data
  mutate(sports_balls = round(sports_balls, digits = 0))
```

Read and clean precipitation data for 2018 and 2019, and combine them.

```{r}
data_2018 = read_excel("./data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", sheet = "2018 Precipitation", range = "A2:B14") %>% 
  janitor::clean_names() %>%
  drop_na(total) %>% 
  mutate(year = "2018")

data_2019 = read_excel("./data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", sheet = "2019 Precipitation", range = "A2:B14") %>% 
  janitor::clean_names() %>%
  drop_na(total) %>% 
  mutate(year = "2019")

data_combine = 
  bind_rows(data_2018, data_2019) %>% 
  mutate(month = month.name[month]) %>% 
  relocate(year, month)
```

Write a paragraph of these data.

* For Mr.Trash Wheel dataset, there are **`r nrow(trash_data)`** rows and **`r ncol(trash_data)`** columns. The number of observations is **`r nrow(trash_data)`**. Variables contain _`r names(trash_data)`_.  Examples of key variables are:
`r knitr::kable(head(select(trash_data, dumpster, month, year, weight_tons, plastic_bottles)))`

* For the combined precipitation dataset of 2018 and 2019, there are **`r nrow(data_combine)`** rows and **`r ncol(data_combine)`** columns. The number of observations is **`r nrow(data_combine)`**. Variables contain _`r names(data_combine)`_. Examples of key variables are:
`r knitr::kable(head(data_combine))`

* The total precipitation in 2018 is **`r sum(pull(filter(data_combine, year == 2018), total))`**. The median number of sports balls in a dumpster in 2019 is **`r median(pull(filter(trash_data, year == 2019),sports_balls))`**.


# Problem 2

First, clean the data in pols-month.csv. 

```{r}
pols_month = read_csv("./data/pols-month.csv") %>% 
  janitor::clean_names() %>% 
  separate(mon, into = c("year", "month", "day")) %>% 
  mutate(month = month.name[as.numeric(month)]) %>% 
  mutate(president = ifelse(prez_gop == 1, "gop", "dem")) %>% 
  select(-day, -prez_gop, -prez_dem) %>% 
  arrange(year, month)
``` 

Second, clean the data in snp.csv.

```{r}
snp = read_csv("./data/snp.csv") %>% 
  janitor::clean_names() %>% 
  separate(date, into = c("month", "day", "year")) %>% 
  mutate(year = paste0(ifelse(as.numeric(year) <= 15, "20", "19"), year)) %>% 
  mutate(month = month.name[as.numeric(month)]) %>%   
  select(-day) %>% 
  relocate(year) %>% 
  arrange(year, month)
```

Third, tidy the unemployment data.

```{r}
unemployment = read_csv("./data/unemployment.csv") %>% 
  pivot_longer(
      Jan:Dec, 
      names_to = "month",
      values_to = "percentage of unemployment"
    ) %>% 
  janitor::clean_names() %>% 
  mutate(
      month = month.name[as.numeric(match(month, month.abb))],
      year = as.character(year)
      ) %>% 
  drop_na(percentage_of_unemployment) %>% 
  arrange(year, month)
```

Fourth, join the datasets.

```{r}
# Join dataset pols_month and snp first.
join_a = left_join(pols_month, snp, by = c("year", "month")) 

# Then join dataset unemployment
join = left_join(join_a, unemployment, by = c("year", "month"))
```

Write a paragraph of these data.

* The dataset `pols_month` gives information about the number of democratic and republican politicians, which contain **`r nrow(pols_month)`** observations and **`r ncol(pols_month)`** variables. Variables contain _`r names(pols_month)`_. The time is from `r min(as.numeric(pull(pols_month, year)))` to `r max(as.numeric(pull(pols_month, year)))`, and the range is `r max(as.numeric(pull(pols_month, year))) - min(as.numeric(pull(pols_month, year)))`.

* The dataset `snp` reveals information about Standard & Poor’s stock market index (S&P), which contain **`r nrow(snp)`** observations and **`r ncol(snp)`** variables. Variables contain _`r names(snp)`_. The time is from `r min(as.numeric(pull(snp, year)))` to `r max(as.numeric(pull(snp, year)))`, and the range is `r max(as.numeric(pull(snp, year))) - min(as.numeric(pull(snp, year)))`.

* The dataset `unemployment` shows percentage of unemployment in a given time, which contain **`r nrow(unemployment)`** observations and **`r ncol(unemployment)`** variables. Variables contain _`r names(unemployment)`_. The time is from `r min(as.numeric(pull(unemployment, year)))` to `r max(as.numeric(pull(unemployment, year)))`, and the range is `r max(as.numeric(pull(unemployment, year))) - min(as.numeric(pull(unemployment, year)))`.


* The dataset `join` combines the above three datasets, and contains **`r nrow(join)`** observations and **`r ncol(join)`** variables.The time is from `r min(as.numeric(pull(join, year)))` to `r max(as.numeric(pull(join, year)))`, and the range is `r max(as.numeric(pull(join, year))) - min(as.numeric(pull(join, year)))`.
Examples of key variables are:
`r knitr::kable(head(join))`

# Problem 3

First, load and tidy the data. 

```{r}
baby = read_csv("./data/Popular_Baby_Names.csv") %>%
  janitor::clean_names() %>% 
  dplyr::distinct() %>% 
  mutate(
    ethnicity = recode(ethnicity,
                       "ASIAN AND PACI" = "ASIAN AND PACIFIC ISLANDER",
                       "BLACK NON HISP" = "BLACK NON HISPANIC", 
                       "WHITE NON HISP" = "WHITE NON HISPANIC"),
    childs_first_name = str_to_title(childs_first_name),
    gender = str_to_upper(gender)
  ) 
```

Produce table showing the rank in popularity of the name “Olivia” as a female baby name over time.

```{r}
Olivia = 
  baby %>% 
  filter(childs_first_name == "Olivia", gender == "FEMALE") %>%
  select(year_of_birth, ethnicity, rank) %>% 
  arrange(year_of_birth,ethnicity) %>% 
  pivot_wider(
    names_from = year_of_birth,
    values_from = rank
  )
```

`r knitr::kable(Olivia)`


Produce a similar table showing the most popular name among male children over time.

```{r}
male_name = 
  baby %>% 
  filter(gender == "MALE", rank == 1) %>%
  select(year_of_birth, ethnicity, childs_first_name) %>% 
  arrange(year_of_birth,ethnicity) %>% 
  pivot_wider(
    names_from = year_of_birth,
    values_from =  childs_first_name
  )
```

`r knitr::kable(male_name)`


Finally, for male, white non-hispanic children born in 2016, produce a scatter plot showing the number of children with a name (y axis) against the rank in popularity of that name (x axis).

```{r}
  baby %>% 
  filter(gender == "MALE", ethnicity == "WHITE NON HISPANIC", year_of_birth == 2016) %>%  
  ggplot(aes(x = rank, y = count)) + 
  geom_point() +
    labs(
      title = "Numer of children with a name vs. Rank in popularity of the name",
      x = "Rank in popularity of the name",
      y = "Number of children with a name"
    )
```

