---
title: "P8105_hw2_xh2470"
author: "Xueqing Huang"
output: github_document
---

```{r setup}
library(tidyverse)
library(readxl)
options(warn = -1)
```

# Problem 1

Read and clean the Mr.Trash Wheel sheet.

```{r}
trash_data = read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = "Mr. Trash Wheel", range = "A2:N408") %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(sports_balls = round(sports_balls, digits = 0))
```

Read and clean precipitation data for 2018 and 2019, and combine them.

```{r}
data_2018 = read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = "2018 Precipitation", range = "A2:B14") %>% 
  janitor::clean_names() %>%
  drop_na(total) %>% 
  mutate(year = "2018")

data_2019 = read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = "2019 Precipitation", range = "A2:B14") %>% 
  janitor::clean_names() %>%
  drop_na(total) %>% 
  mutate(year = "2019")

data_combine = 
  bind_rows(data_2018, data_2019) %>% 
  mutate(month = month.name[month])
```

Write a paragraph of these data.

```{r filter data}
trash_data_filter = filter(trash_data, year == "2017")

# The total precipitation in 2018 is `r sum(data_2018, total)`.
#The median number o sports balls in a dumpster in 2017 is `r median(trash_data_filter, sports_balls)`
```

For Mr.Trash Wheel dataset, there are `r nrow(trash_data)` rows and `r ncol(trash_data)` columns. Variables contain _`r names(trash_data)`_.
For the combined precipitation dataset of 2018 and 2019, there are `r nrow(data_combine)` rows and `r ncol(data_combine)` columns. Variables contain _`r names(data_combine)`_. 




